FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-12 NewVersion/sections/Conclusion.tex
Example in intro

This paper proposes \MIM{} as a new multiple inheritance model for unintentional method conflicts.
Existing approaches including static dispatch and dynamic dispatch have to compromise between code reuse
and type safety, whereas \MIM{} uses a different solution called hierarchical dispatch to obtain both.
Such an approach not only offers great code reuse like dynamic dispatch but also ensures unambiguity by
our algorithm for method resolution. This paper also introduces hierarchical overriding that refines branches individually.
\MIM{} is formalized with a basis on Featherweight Java and supported
by a few theorems. The prototype is implemented in Scala as a simple interpreter.

Our model can certainly be improved at some aspects. We did not formalize fields as in Featherweight Java for simplicity.
And we did not formalize other common features including
method overloading, casts, covariant return types, and so on, some of which are orthogonal in the design space. Moreover, we restrict that hierarchical overriding can only work on original methods. Potentially a looser condition can better support encapsulation and modularity with respect to code design.

>>>>

This paper proposes \MIM{} as a new multiple inheritance model for
unintentional method conflicts. Previous approaches 
either do not support unintentional method conflicts, have to
compromise between code reuse and type safety, or do not fully support
overriding in the presence of unintentional conflicts. To deal with unintentional method conflicts we
introduce two key mechanisms: hierarchical dispatching and
hierarchical overriding. Hierarchical dispatching is inspired by the
mechanisms in C++. We provide a minimal formal model of hierarchical
dispatching in \MIM{}. Such algorithm makes use of both dynamic type
information and path information from either upcasts or parameters'
information. Such an approach not only offers great code reuse like
dynamic dispatch but also ensures unambiguity by our algorithm for
method resolution. Additionally we introduce \emph{hierarchical
  overriding} to allow conflicting methods in different branches to be
individually overriden. \MIM{} is formalized following the style of
Featherweight Java and type-soundness is proved. A prototype language is
implemented in Scala as a simple interpreter.

Our model can certainly be improved at some aspects. We do not
formalize fields as in Featherweight Java for the reason of
simplicity\bruno{I'm not sure that simplicity is a good excuse; 
following the trait model may be a better excuse.}. And we do not formalize other common features including
method overloading, casts, covariant return types, and so on, some of
which are orthogonal in the design space. Moreover, we restrict that
hierarchical overriding can only work on original methods. Potentially
a looser condition can better support encapsulation and modularity
with respect to code design as discussed in
Section~\ref{sec:discussion} and could be a good future work.

### 2017-10-12 NewVersion/sections/Introduction.tex
Example in intro

\yanlin{
TODOs:
\begin{itemize}
\item UML Diagram for full solution in introduction with informal explanation
\item Move Figure 5 to end of Section 2 together with explanation
\item Improve Figure 5 with additional mbody calls
\item Improve Figure 5 with visual information that signals that in examples d) and e) then D does not type-check
\item Make names of classes more consistent in Figure 5
\item Add example (f) : A m      B m   C m   D  m  This would compile, note that if it was BArrowA m and/or CArrowA m it would also compile and D would have the same observable behaviour
\item remove S1 “Unfortunately, such an issue ...”
\item S2, should remove the discussion about JFrmae. all return type void.
\item S2, mention DrawableDeck in P2.3 is a new version of DrawableDeck before.
\item S2, “Ideally, we want that invocation to be dynamically dispatched.” --> “In FHJ, this invocation is dynamically dispatched.“. and revise the following paragraph on “we may have to copy the shuffleAndDraw code into SafeDeck, so as to be adaptive to the new draw. “.
\item S2.2 revise paragraph “Implicit and Explicit Upcasts “.
\item S2.3 (Maybe) put “Potential solutions/workarounds in existing languages” after “FHJ solution”.
\item S2.3 (Maybe) add “redraw()” method in Drawable. and add another method call similar to shuffleAndDraw()
\item Talk about Fig5 earlier. Move from S4.1 to S2.3 (a peek of ..)
\item The keywords in code are barely distinghuishable from other code,  make them bolder/darker

@@

This paper proposes two mechanisms to deal with unintentional method
conflicts: \textit{hierarchical dispatching} and \emph{hierarchical
  overriding}. Hierarchical dispatching is inspired by the mechanisms in C++ and
provides an approach
to method dispatching, which combines static and dynamic
information. Using hierarchical dispatching, the method binder will look
at both the \emph{static type} and the \emph{dynamic type} of the
receiver during runtime. When there are multiple branches that cause
unintentional conflicts, the static type can specify one branch among
them for unambiguity, and the dynamic type helps to find the most
specific implementation. In that case, both unambiguity and
extensibility are preserved. The main novelty over existing work is 
the formalization of the essence of a hierarchical dispatching
algorithm, which (as far as we know) has not been formalized before. 
Furthermore, \textit{hierarchical overriding} allows method overriding to be applied
only to one branch of the class hierarchy. Hierarchical overriding 
adds expressive power that is not available in languages such as 
C++ or C\#. To present both ideas, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness.

>>>>

\yanlin{
TODOs:
\begin{itemize}
\item UML Diagram for full solution in introduction with informal explanation
\item Move Figure 5 to end of Section 2 together with explanation
\item Improve Figure 5 with additional mbody calls
\item Improve Figure 5 with visual information that signals that in examples d) and e) then D does not type-check
\item Make names of classes more consistent in Figure 5
\item Add example (f) : A m      B m   C m   D  m  This would compile, note that if it was BArrowA m and/or CArrowA m it would also compile and D would have the same observable behaviour
\item remove S1 “Unfortunately, such an issue ...”
\item S2, should remove the discussion about JFrmae. all return type void.
\item S2, mention DrawableDeck in P2.3 is a new version of DrawableDeck before.
\item S2, “Ideally, we want that invocation to be dynamically dispatched.” --> “In FHJ, this invocation is dynamically dispatched.“. and revise the following paragraph on “we may have to copy the shuffleAndDraw code into SafeDeck, so as to be adaptive to the new draw. “.
\item S2.2 revise paragraph “Implicit and Explicit Upcasts “.
\item S2.3 (Maybe) put “Potential solutions/workarounds in existing languages” after “FHJ solution”.
\item S2.3 (Maybe) add “redraw()” method in Drawable. and add another method call similar to shuffleAndDraw()
\item Talk about Fig5 earlier. Move from S4.1 to S2.3 (a peek of ..)
%\item The keywords in code are barely distinghuishable from other code,  make them bolder/darker

@@

This paper proposes two mechanisms to deal with unintentional method
conflicts: \textit{hierarchical dispatching} and \emph{hierarchical
  overriding}. Hierarchical dispatching is inspired by the mechanisms in C++ and
provides an approach
to method dispatching, which combines static and dynamic
information. Using hierarchical dispatching, the method binder will look
at both the \emph{static type} and the \emph{dynamic type} of the
receiver during runtime. When there are multiple branches that cause
unintentional conflicts, the static type can specify one branch among
them for unambiguity, and the dynamic type helps to find the most
specific implementation. In that case, both unambiguity and
extensibility are preserved. The main novelty over existing work is 
the formalization of the essence of a hierarchical dispatching
algorithm, which (as far as we know) has not been formalized before.

\begin{figure}
\center  \includegraphics[height=4cm]{pics/DrawableSafeDeck3.pdf}
\caption{\lstinline|DrawableSafeDeck|: an illustration of hierarchical overriding.}
\label{fig:main}
\end{figure}

\textit{Hierarchical overriding} is a novel language
mechanism that allows method overriding to be applied
only to one branch of the class hierarchy. Hierarchical overriding 
adds expressive power that is not available in languages such as 
C++ or C\#. Hierarchical overriding allows overriding to work 
for classes with multiple (conflicting) methods sharing the 
same names and signatures. An example is presented in 
Figure~\ref{fig:main}. In this example there are 4 classes/interfaces.
Two classes model a deck of cards (\lstinline|Deck|) and a class for
drawable widgets (\lstinline|Drawable|). 
A class \lstinline|SafeDeck| adds functionality to check whether a 
the deck is empty and to prevent drawing a card from an empty deck. 
The interesting class is \lstinline|DrawableSafeDeck|, which inherits 
from both \lstinline|SafeDeck| and \lstinline|Drawable|. Hierchical
overriding is used in \lstinline|DrawableSafeDeck| to keep two separate \lstinline|draw|
methods for each parent, but overriding \emph{only} the \lstinline|draw|
method coming from \lstinline|Drawable|, in order to draw a widget
with a deck of cards. Note that hierchical overriding is denoted in the UML
diagram with the notation
\lstinline|draw()|$\shneg$\lstinline|Drawable|. In this case the
notation expresses that the \lstinline|draw| method from \lstinline|Drawable|
is overriden. Although in this example only one of the
\lstinline|draw| methods is overriden (and the other is simply inherited), hierarchical overriding
supports multiple conflicting methods to be independently overriden as well.

To present hierarchical overriding and dispatching, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness.
